\large UNet-HCRF Integration for Sequence Labeling on EEG Data

### 2024-12-09 manuscript/slides.tex
edit intro; add framework

\begin{frame}{Introduction}
\begin{itemize}
\setlength{\itemsep}{1.5em}
\item Sequence labeling tasks are crucial for healthcare, where accurate and
timely data classification can significantly impact patient outcomes.
\item Deep learning has recently emerged as a powerful tool for sequence
labeling.
\item Conditional random fields (CRFs) have been employed for modeling label
dependencies, demonstrating promising results.
\end{itemize}
\end{frame}

@@

\begin{frame}{Variational Inference of HCRF}
\begin{small}
\begin{itemize}
\item The mean-field varepsilon family approximation of
$\log P(\hb \mid y,\xb,\theta)$ is expressed as
$\QQ(h) = \prod_{i=1}^N q_i(h_i)$.
\item The evidence lower bound (ELBO) under the mean-field family is:
\begin{equation*}
\begin{split}
& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \right] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
\varpi).
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Variational Inference of HCRF (cont.)}
\begin{itemize}
\item By the coordinate ascent variational inference(CAVI) algorithm, fixing
$q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
$\mathrm{ELBO}(q_i(h_i))$ is:
\begin{equation*}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
\end{equation*}
\item The explicit form of $q_i(l)$ is derived as follows:
\begin{equation*}
\begin{split}
& q_i(l) = \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
& \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\right)
\right]
\Bigg\},
\end{split}
\end{equation*}
where $Z_i$ is the normalization constant.
\end{itemize}
\end{frame}

@@

\begin{frame}{Harmful Brain Activity Classification Dataset}
\begin{itemize}
\item The dataset contains 50-second long electroencephalography (EEG) samples
covering a 10-minute window as well as spectrograms matching each sample.
\bigskip
\item The EEG samples were labeled by expert annotators to determine which of
the six features each spectrogram belonged to.
\bigskip
\item The longer 10-minute window allows for the model to pick up on further
insights regarding the nature of the brain activity.
\end{itemize}
\end{frame}

>>>>

\begin{frame}{Introduction}
\begin{itemize}
\setlength{\itemsep}{1.5em}
\item Electroencephalography (EEG) pattern recognition tasks are crucial for
healthcare, where accurate and timely data classification can significantly
impact patient outcomes.
\item Deep learning has recently emerged as a powerful tool for sequence
labeling.
\item Conditional random fields (CRFs) have been employed for modeling label
dependencies, demonstrating promising results.
\end{itemize}
\end{frame}

\begin{frame}{Framework}
\begin{figure}[tbp]
\centering
\includegraphics[width=\textwidth]{model}
\end{figure}
\end{frame}

@@

\begin{frame}{Variational Inference of HCRF}
\begin{small}
\begin{itemize}
\item The mean-field varepsilon family approximation of
$\log P(\hb \mid y,\xb,\theta)$ is expressed as
$\QQ(h) = \prod_{i=1}^N q_i(h_i)$.
\item The evidence lower bound (ELBO) under the mean-field family is:
\begin{equation*}
\begin{split}
& \mathrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \right] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
\varpi).
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Variational Inference of HCRF (cont.)}
\begin{itemize}
\item By the coordinate ascent variational inference (CAVI) algorithm, fixing
$q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
$\mathrm{ELBO}(q_i(h_i))$ is:
\begin{equation*}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
\end{equation*}
\item The explicit form of $q_i(l)$ is derived as follows:
\begin{equation*}
\begin{split}
& q_i(l) = \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
& \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\right)
\right]
\Bigg\},
\end{split}
\end{equation*}
where $Z_i$ is the normalization constant.
\end{itemize}
\end{frame}

@@

\begin{frame}{Harmful Brain Activity Classification Dataset}
\begin{itemize}
\item The dataset contains 50-second long EEG samples covering a 10-minute
window as well as spectrograms matching each sample.
\bigskip
\item The EEG samples were labeled by expert annotators to determine which of
the six features each spectrogram belonged to.
\bigskip
\item The longer 10-minute window allows for the model to pick up on further
insights regarding the nature of the brain activity.
\end{itemize}
\end{frame}
